---
title: "Cell Type Density"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(spatstat)
library(sp)
library(reshape2)
library(GWmodel)
library(ggplot2)
library(stringr)
library('plot.matrix')
```

To better compare across samples, we want to make sure squares are the same size, so we'll set a square size and tell the algorithm 

```{r}
# parameters for density estimation
thres = 1e-4
square_size=10
mincount=5

# column names
X_col = "X.X"
Y_col = "Y.Y"
pheno_col = "ClusterName"

# cell type we're calculating density  for
celltype = "tumor cells"

```

```{r}
XYdata <- read.table("CODEX Data/1.csv", header = TRUE, sep = ",", fill = TRUE, row.names=NULL)
head(XYdata)
```

To calculate the number of grids we need, we identify the range for X and Y coordinates, then calculate the appopriate step size.

```{r}
xmin = min(XYdata[,X_col])
xmax = max(XYdata[,X_col])
ymin = min(XYdata[,Y_col])
ymax = max(XYdata[,Y_col])
  
xstep = ceiling(xmax/square_size)
ystep = ceiling(ymax/square_size)
kde_grid_size = c(ystep,xstep )
```

Next, we create our point pattern process as calculate density with respect to our cell type of interest.
```{r}
dataPoints=XYdata[which(XYdata[,pheno_col] == celltype),]
W = owin(c(xmin,xmax),c(ymin,ymax))
ref_ppp = as.ppp(dataPoints[,c(X_col, Y_col)],W)
kde_ref = density(ref_ppp, positive = T, dimyx=kde_grid_size)

```

```{r}
kde_ref$v[1:5,1:5]
```

The original output is an nxn matrix, where n is the number of grids. This is more of an "image" style format.

```{r}
# since we loaded the plot.matrix package, this will be used instead of the base R plotting tool
plot(kde_ref$v, border=NA)
```

But we also might want to melt this matrix, i.e. convert it to a matrix/df where each row contains an X/Y coordinate and then the density value at that coordinate. This format is really helpful if you want to calculate & compare multiple densities at the same coordinate.

```{r}
kde_values = kde_ref$v
row.names(kde_values) = kde_ref$yrow
colnames(kde_values) = kde_ref$xcol
    
values = melt(kde_values)
colnames(values) <- c("Y", "X", celltype)
```


```{r}
head(values)
```
The "melted" format is also the format that works with ggplot2.

```{r}
ggplot() + geom_tile(data=values, aes(x=X, y=Y, fill=`tumor cells`)) + geom_point(data=XYdata, aes_string(x=X_col, y=Y_col))
```

This can be easily adapted for any other kind of density-esque calculations, like [GWR](https://cran.r-project.org/web/packages/spgwr/vignettes/GWR.pdf).








